\MakeUppercase{The Structure of $k$-Planar Graphs}

### 2020-03-25 k-planar.tex
Pffft

 Otherwise, $V(X)\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

@@

\cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.

\begin{cor}
Every $(g,\delta)$-string graph has $p$-centered chromatic number
\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
\tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
and every plane $(0,\delta)$-string graph has $p$-centered chromatic number
\[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
\tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
\end{cor}
%\label{p-centered}
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.

@@

\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number of $(g,\delta)$-string graphs, but the bound is weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ by a result of \citet{DJKW16}.

@@

\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in \cref{Introduction}.

@@

\begin{cor}
\label{PowerCor}
For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
\begin{compactitem}
%%%%%%
\item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
%%%%%%
\item $G^k$ has queue-number
\[\qn(G^k) \leq 3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
%If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
%%%%%%
\item $G^k$ has  $p$-centered chromatic number
\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1) \tbinom{k+4}{4}
\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace,\]
%\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
\item $G^k$ has  non-repetitive chromatic number
\[ \pi(G^k) \leq  \max\{2g,3\} (2k^4+6k^2) \Delta^{k} \, 4^{\binom{k+4}{4}}\enspace.\]
%If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$,  then $\pi(G)\le \ell\, 4^{t+1}$.
\end{compactitem}
\end{cor}

This result is the first constant upper bound on the queue-number and $p$-centered chromatic number of bounded powers of graphs with bounded degree and bounded Euler genus.  For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$. \cref{PowerCor} improves upon this bound when $k\ll\Delta$ and $G$ has Euler genus $g$.

@@

We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n$ vertices has at least $kn/2$ edges and at most $kn$ edges.  For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.

>>>>

 Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

@@

% \snote{Deleted results on p-centered colouring since the bounded-degree results by \citet{debski.felsner.ea:improved} give better dependence on $\delta$.}
%
%
% \cref{p-centered,p-centered-treewidth,StringPartition} imply the first known bounds on the $p$-centered chromatic number of map graphs that are polynomial in $p$.
%
% \begin{cor}
% Every $(g,\delta)$-string graph has $p$-centered chromatic number
% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
% \tbinom{ \delta+5}{4} \binom{p+\binom{ \delta+5}{4}-1}{\binom{ \delta+5}{4}-1}\enspace, \]
% and every plane $(0,\delta)$-string graph has $p$-centered chromatic number
% \[ \chi_p(G)\le \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4) (p+1)
% \tbinom{ \delta+4}{3}\binom{p+\binom{ \delta+4}{3}-1}{\binom{ \delta+4}{3}-1}\enspace.\]
% \end{cor}
%\label{p-centered}
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.

@@

\cref{non-repetitive,StringPartition} also give bounds on the non-repetitive chromatic number and the $p$-centered colouring number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{debski.felsner.ea:improved}, respectively.

@@

\cref{PlanarPartition}(b), \cref{SurfacePartition}(b), \cref{modern-major-general,PowerShortcut} imply the following result, which with \cref{PartitionProduct} implies \cref{kPowerBasic} in the introduction.

@@

\begin{cor}
\label{PowerCor}
For all integers $k\geq 1$ and $g\geq 0$, if $G$ is a graph with Euler genus $g$ and maximum degree $\Delta$, then:
\begin{compactitem}
%%%%%%
\item $G^k$ is $(g, 2k(k+1)\Delta^{k} )$-planar,
%%%%%%
\item $G^k$ has queue-number
\[\qn(G^k) \leq 3\,\max\{2g,3\} (2k^4+6k^2) \,\Delta^{k}\, 2^{\binom{k+4}{4}-1},\]
%If a graph $G$ has a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t$, then $\qn(G)\le 3 \ell (2^t-1) + \floor{\tfrac{3}{2}\ell} \le 3 \ell (2^t-\frac12 )$.
%%%%%%
\item $G^k$ has  $p$-centered chromatic number
\[\chi_p(G^k) \leq \max\{2g,3\}(2k^4+6k^2) \Delta^{k}  (p+1)
\binom{p+\binom{k+4}{4}-1}{\binom{k+4}{4}-1}\enspace,\]
%\label{p-centered} Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
\item $G^k$ has  non-repetitive chromatic number
\[ \pi(G^k) \leq  \max\{2g,3\} (2k^4+6k^2) \Delta^{k} \, 4^{\binom{k+4}{4}}\enspace.\]
%If a graph $G$ has an $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$,  then $\pi(G)\le \ell\, 4^{t+1}$.
\end{compactitem}
\end{cor}

This result is the first constant upper bound on the queue-number of bounded powers of graphs with bounded degree and bounded Euler genus.  For every graph $G$, since $G^k$ has maximum degree at most $\Delta^k$, a result of \citet{DJKW16} implies that $\pi(G^k) \leq (1+o(1))\Delta^{2k}$. \cref{PowerCor} improves upon this bound when $k\ll\Delta$ and $G$ has Euler genus $g$.  Similarly, a result of \citet{debski.felsner.ea:improved} implies that $\chi_p(G^k)\le 1024p\Delta^{2k}$ and \cref{PowerCor} improves upon this bound when $p,k\ll\Delta$ and $G$ has Euler genus $g$.

@@

We note that \cref{nearest-neighbour} is tight up to the leading constant:  Every $k$-nearest neighbour graph on $n\ge k+1$ vertices has at least $kn/2$ edges and at most $kn$ edges.  For $k\ge 7$, the Crossing Lemma~\citep{ajtai.chvatal.ea:crossing-free,leighton:complexity} implies that the total number of crossings is therefore $\Omega(k^3n)$ so that the average number of crossings per edge is $\Omega(k^2)$.
